paper3

### 2022-08-26 paper3.Rmd
grid 1 km

X_train <- train_set %>% 
  dplyr::select(Scenicness = Scenicness.median_of_bayescore,
                Wildness = Wildness.median, x_coord, y_coord)
 X_test <-  test_set %>% 
   dplyr::select(Scenicness = Scenicness.median_of_bayescore,
                 Wildness = Wildness.median, x_coord, y_coord)

@@

library(reticulate)
# -- creates an interactive Python console, REPL (Read, Eval, Print Loop), within R.
repl_python()

# reticulate::py_config() # why numpy_version couldn't be found?
# use_python("Z:\\R\\R projects\\LCA-GB\\my_env_Python_3.9.13\\Scripts\\python.exe")
# 
# # use_virtualenv("Z:\\R\\R projects\\LCA-GB\\my_env_Python_3.9.13", required = TRUE)
# 
# install.packages("usethis")
# usethis::edit_r_environ()
# usethis::edit_r_profile()
```
```{python}
# -- load in the train/test sets
X_train = r.X_train
y_train = r.y_train
 X_test = r.X_test
 y_test = r.y_test

@@

model = xgb.train(space_eval(param_space, best_params),
                  dtrain,
                  num_boost_round = 500,
                  evals = [(dtrain, 'train'), (dtest, 'test')],
                  early_stopping_rounds = 250)

# model = xgb.train(params,
#                   dtrain,
#                   num_boost_round = 500,
#                   evals = [(dtrain, 'train'), (dtest, 'test')],
#                   verbose_eval = 100,
#                   early_stopping_rounds = 251)

@@

```{python plot}
X =  r.landmap.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x', 'y']]
# -- play shap
explainer = shap.TreeExplainer(model)
# X = landmap.loc[:, ['UID', 'Scenicness.median_of_bayescore', 'Wildness.median', 'x', 'y']]
# The approximate argument set to True to fix some of the shap_values with NaN value
# shape of shap_values is : (4, 1321, 4) means 4 classes, 1321 rows(observations), each 4 columns(features).
shap_values = explainer.shap_values(X, approximate = True)

@@

# In the case of matplotlib version 3.5 above, the adjustment of colorbar legend was referred to https://github.com/slundberg/shap/issues/2493
# summary plot for "CLASS 0" (Low scenic quality) 
shap.summary_plot(shap_values[0],
                  X_train.values,
                  feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], #X_train.columns,
                  show = False,
                  # plot_type = 'compact_dot', # making the dots disappear
                  plot_size = (50, 6))
plt.gcf().axes[-1].set_aspect('auto')
plt.tight_layout()
plt.gcf().axes[-1].set_box_aspect(40) # smaller "box_aspect" value to make colorbar thicker
plt.show(); plt.close()
# plt.savefig('fig_tes1.png', bbox_inches='tight',dpi=300); plt.close()

# X_names = ['pct_18_34','pct_white','pct_bach',
#           'pct_no_car','log_pop_den','job_entropy','network_den',
#           'ave_trip_dist','pct_share','x_coord','y_coord']
# 
# fig, ax = plt.subplots(figsize=(8, 6),dpi=160)
# 
# shap.summary_plot(shap_interaction_values, X_coords, max_display=15, 
#                   feature_names = X_names,
#                   plot_type="compact_dot")
```
## Spatial patterns of SHAP values
```{python}
# landmap = r.landmap
X_vars = ['Scenicness','Wildness', 'x', 'y']

@@

for j in range(len(X_vars)):
  j = 1
    ax = ax.ravel()
    X.plot(ax=ax[j],column = shap_values.values[:,j],legend=True,
                 vmin=-0.8,vmax=0.8,cmap=shap.plots.colors.red_white_blue)

    ax[j].set_title("SHAP for\n" + X_vars[j],fontsize=10)

#fig.delaxes(ax[-1])

```{r}
shap_values = py$shap_values
shap_Low = data.frame(shap_values[[1]]) %>% transmute(SHAP_Scenicness_Low = X1, 
                                                      SHAP_Wildness_Low = X2,
                                                      SHAP_Location_Low = X3 + X4) 
shap_Moderate = data.frame(shap_values[[2]]) %>% transmute(SHAP_Scenicness_Moderate = X1,
                                                           SHAP_Wildness_Moderate = X2,
                                                           SHAP_Location_Moderate= X3 + X4)
shap_High = data.frame(shap_values[[3]]) %>% transmute(SHAP_Scenicness_High = X1,
                                                       SHAP_Wildness_High = X2,
                                                       SHAP_Location_High = X3 + X4)
shap_Outstanding = data.frame(shap_values[[4]]) %>% transmute(SHAP_Scenicness_Outstanding = X1,
                                                              SHAP_Wildness_Outstanding = X2,
                                                              SHAP_Location_Outstanding = X3 + X4)

@@

## map together
setwd("Z:\\R\\R projects\\LCA-GB\\figures")
png(filename = "SHAP_values.png", w = 10*2, h = 6*2, units = "in", res = 300)
par(mar = c(0,0,0,0))
multiplot(list(f[1], f[2], f[3], f[4], 
               f[5], f[6], f[7], f[8], 
               f[9], f[10], f[11], f[12]), cols = 4)
dev.off()
```

>>>>

X_train <- train_set %>% 
  dplyr::select(Scenicness = Scenicness.weighted_bayescore,
                Wildness = Wildness.median, x_coord, y_coord)
 X_test <-  test_set %>% 
   dplyr::select(Scenicness = Scenicness.weighted_bayescore,
                 Wildness = Wildness.median, x_coord, y_coord)

@@

library(reticulate)
# -- creates an interactive Python console, REPL (Read, Eval, Print Loop), within R.
repl_python() 
```

```{python}
# -- load in the train/test sets
X_train = r.X_train
y_train = r.y_train
 X_test = r.X_test
 y_test = r.y_test

@@

# model = xgb.train(space_eval(param_space, best_params),
#                   dtrain,
#                   num_boost_round = 500,
#                   evals = [(dtrain, 'train'), (dtest, 'test')],
#                   early_stopping_rounds = 250)

model = xgb.train(params,
                  dtrain,
                  num_boost_round = 500,
                  evals = [(dtrain, 'train'), (dtest, 'test')],
                  verbose_eval = 100,
                  early_stopping_rounds = 251)

@@

```{python plot}
X =  r.landmap.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x_coord', 'y_coord']]
# -- play shap
explainer = shap.TreeExplainer(model)

# The approximate argument set to True to fix some of the shap_values with NaN value
# shape of shap_values is : (4, 1321, 4) means 4 classes, 1321 rows(observations), each 4 columns(features).
shap_values = explainer.shap_values(X, approximate = True)

@@

# In the case of matplotlib version 3.5 above, the adjustment of colorbar legend was referred to https://stackoverflow.com/questions/70461753/shap-the-color-bar-is-not-displayed-in-the-summary-plot/
# summary plot for "CLASS 0" (Low scenic quality)

# landmap = r.landmap
X_vars = ['Scenicness','Wildness', 'x_coord', 'y_coord']

import matplotlib.pyplot as plt

@@

for j in range(len(shap_values)):
    ax = ax.ravel()
    # X.plot(ax=ax[j],column = shap_values.values[:,j],legend=True,
    #              vmin=-0.8,vmax=0.8,cmap=shap.plots.colors.red_white_blue)
    fig[j] = shap.summary_plot(shap_values[1],
                              X.values,
                              feature_names = X_vars,
                              show = False),
                              # plot_type = 'compact_dot', # making the dots disappear
                              plot_size = (50, 6))
             plt.gcf().axes[-1].set_aspect('auto')
             plt.tight_layout()
             plt.gcf().axes[-1].set_box_aspect(40)
    ax[j].set_title("SHAP for\n" + X_vars[j],fontsize=10)

fig.delaxes(ax[-1])

shap.summary_plot(shap_values[1],
                  X.values,
                  feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], show = False),
                  # plot_type = 'compact_dot', # making the dots disappear
                  plot_size = (50, 6))
plt.gcf().axes[-1].set_aspect('auto')
plt.tight_layout()
plt.gcf().axes[-1].set_box_aspect(40) # smaller "box_aspect" value to make colorbar thicker
plt.show(); plt.close()
# plt.savefig('fig_tes1.png', bbox_inches='tight',dpi=300); plt.close()
```
## Spatial patterns of SHAP values
```{r}
shap_values = py$shap_values
shap_Low = data.frame(shap_values[[1]]) %>% transmute(SHAP_Scenicness_Low = X1, 
                                                      SHAP_Wildness_Low = X2,
                                                      SHAP_Location_Low = X3 + X4) 
shap_Moderate = data.frame(shap_values[[2]]) %>% transmute(SHAP_Scenicness_Moderate = X1,
                                                           SHAP_Wildness_Moderate = X2,
                                                           SHAP_Location_Moderate= X3 + X4)
shap_High = data.frame(shap_values[[3]]) %>% transmute(SHAP_Scenicness_High = X1,
                                                       SHAP_Wildness_High = X2,
                                                       SHAP_Location_High = X3 + X4)
shap_Outstanding = data.frame(shap_values[[4]]) %>% transmute(SHAP_Scenicness_Outstanding = X1,
                                                              SHAP_Wildness_Outstanding = X2,
                                                              SHAP_Location_Outstanding = X3 + X4)

@@

## map together
setwd("Z:\\R\\R projects\\LCA-GB\\figures")
png(filename = "SHAP_values.png", w = 10*2, h = 6*2, units = "in", res = 300)
par(mar = c(0,0,0,0))
multiplot(list(f[1], f[2], f[3], f[4], 
               f[5], f[6], f[7], f[8], 
               f[9], f[10], f[11], f[12]), cols = 4)
dev.off()
```
# Change of support (to 1 km grid)
```{r}
# -- download the administrative region for Great Britain from GADM
UK <- raster::getData('GADM', country = 'GBR', level = 0) |>
  as('sf') |>
  # filter(!NAME_1=='Northern Ireland') |>
  st_transform(crs = 27700)
plot(UK)

# -- download OSGB 1 km grid from github and covert its coordinate system to 27700
grid_1km <- "https://raw.githubusercontent.com/charlesroper/OSGB_Grids/master/GeoJSON/OSGB_Grid_1km.geojson" |>
  geojsonio::geojson_read(what = "sp") |>
  as("sf") |>
  st_transform(crs = 27700)

st_crs(grid_1km) == st_crs(sc)

# -- spatial join OSGB 1 km grid with Scenic-Or-Not data
test <- st_join(grid_1km, sc)

test1 <- test |>
  aggregate(cbind(ID,
                  `Geograph URI`,
                  Votes,
                  Number,
                  Mean,
                  Median,
                  BayeScore,
                  Entropy,
                  Variance)~PLAN_NO, data = ., paste0, collapse = ",") %>%
  merge(LANDMAP, ., by = "UID", all = T) %>%
  as_tibble()

plot(UK['NAME_0']) +
  plot(test[is.na(test$ID),'Mean'])

grid_1km <- st_join(grid_1km, sc, join = st_within, left = TRUE) |>
  dplyr::select(PLAN_NO, 
                `Scenic-Or-Not ID` = ID,
                `Geograph URI`,
                Votes,
                Number,
                Mean,
                Median,
                BayeScore,
                Entropy,
                Variance)

grid_1km <- st_transform(grid_1km, crs = st_crs(Wildness)) %>%
  exact_extract(Wildness, ., 
                fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), 
                quantiles = c(0.25, 0.75)) |>
  mutate(Wildness.mean = mean,
         Wildness.median = median,
         Wildness.iqr = q75 - q25,
         Wildness.cv = coefficient_of_variation,
         Wildness.variance = variance) |>
  dplyr::select(-c(mean, median, q25, q75, coefficient_of_variation, variance)) |>
  bind_cols(grid_1km, .) |>
  st_transform(crs = 27700)
